---
title: "Sobol Uncertainty Partitioning"
author: "Michael Dietze"
date: "3/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Context

Code to experiment with translating one-at-a-time uncertainty partitioning into a global uncertainty partitioning. Will borrow the extended logistic example from EF_Activities Ch 11 
```{r, echo=FALSE}
EFA = "~/git/EF_Activities/data/" ## local path to EF_Activities data
load(file.path(EFA,"Ch11_UA.RData"))
library(ecoforecastR)
```

## Settings and Functions
```{r}
### settings
s <- 6             ## Focal site for forward simulation
Nmc = 1000         ## set number of Monte Carlo draws
ylim = c(100,700)  ## set Y range on plot
N.cols <- c("black","red","green","blue","orange") ## set colors
trans <- 0.8       ## set transparancy
time = 1:(NT*2)    ## total time
time1 = 1:NT       ## calibration period
time2 = time1+NT   ## forecast period
ci <- apply(as.matrix(out$predict),2,quantile,c(0.025,0.5,0.975)) ## hindcast CI

plot.run <- function(){
  sel = seq(s,ncol(ci),by=NS)
  plot(time,time,type='n',ylim=ylim,ylab="N")
  ecoforecastR::ciEnvelope(time1,ci[1,sel],ci[3,sel],col=col.alpha("lightBlue",0.6))
  lines(time1,ci[2,sel],col="blue")
  points(time1,No[s,])
}

##` @param IC    Initial Conditions
##` @param X     Drivers/covariates
##` @param theta Parameters
##` @param alpha Random effects
##` @param Q     Process error parameter (default = 0 for deterministic runs)
##` @param zp    Process error random draws
##` @param n     Size of Monte Carlo ensemble
forecastN <- function(IC,X,theta,alpha,Q=0,zp=0,n=Nmc){
  r = theta[,"r_global"]
  Kg = theta[,"K_global"]
  beta = theta[,"beta"]
  ppt = X
  N <- matrix(NA,n,NT)  ## storage
  Nprev <- IC           ## initialize
  for(t in 1:NT){
    K = pmax(1,Kg + alpha + beta*log(ppt[,t]/800))  ## calculate carrying capacity
    mu = log(pmax(1,Nprev + r*Nprev*(1-Nprev/K)))   ## calculate mean
#    N[,t] <- rlnorm(n,mu,Q)                         ## predict next step
    N[,t] <- exp(mu+Q*zp)
    Nprev <- N[,t]                                  ## update IC
  }
  return(N)
}
```

## Prep Monte Carlo samples
```{r}
## prep MC sample
params <- as.matrix(out$params)
ICfull <- as.matrix(out$predict)

prow = sample.int(nrow(params),Nmc,replace=TRUE)  ## parameter & IC rows to sample
IC  <- ICfull[prow,"N[6,30]"]
theta <- params[prow,c("r_global","K_global","beta")]
alpha <- params[prow,"alpha_site[6]"]
Qmc <- 1/sqrt(params[prow,"Q"])  ## convert from precision to standard deviation

drow = sample.int(nrow(ppt_ensemble),Nmc,replace=TRUE) ## driver rows to sample
X = ppt_ensemble[drow,]

zp = rnorm(Nmc) ## Z-scores for process error
```


## Default ensemble forecast
```{r}
## run ensemble forecast
Nf <- forecastN(IC    = IC,  ## sample IC
                X     = X,   ## Sample drivers
                theta = theta, ## sample params
                alpha = alpha, ## sample RE
                Q     = Qmc,
                zp    = zp
#                n     = Nmc
                )
Nf.ci = apply(Nf,2,quantile,c(0.025,0.5,0.975))

plot.run()
ecoforecastR::ciEnvelope(time2,Nf.ci[1,],Nf.ci[3,],col=col.alpha(N.cols[4],trans))
```
## prep Monte Carlo set B
```{r}
rowAvail = seq_len(nrow(params))[-prow] ## set of rows not yet sampled
prowB = sample(rowAvail,Nmc,replace=TRUE)
ICB  <- ICfull[prowB,"N[6,30]"]
thetaB <- params[prowB,c("r_global","K_global","beta")]
alphaB <- params[prowB,"alpha_site[6]"]
QmcB <- 1/sqrt(params[prowB,"Q"])  ## convert from precision to standard deviation

drowB = sample.int(nrow(ppt_ensemble),2*Nmc,replace=TRUE) ## oversample
drowB = drowB[-which(drowB == drow)][1:Nmc] ## remove duplicate pairs
  ## note: removing duplicates will overestimate var, but worried about zero var terms
XB = ppt_ensemble[drowB,]

zpB = rnorm(Nmc)
```

## Initial Sobal at CATEGORY level, d=6
```{r}
## Flip A and B to match Saltelli definition of Ci = all B except Ai
A = forecastN(ICB,XB,thetaB,alphaB,QmcB,zpB,Nmc)         # A
B = Nf                                                   # B

## Saltelli eq 4.6: f0 = E(Y)
mu = matrix(apply(Nf,2,mean,na.rm=TRUE),nrow(Nf),ncol(Nf),byrow = TRUE)

## Saltelli eq 4.7: fi = E(Y|Xi)-E(Y)
B = B - mu                                               # B
A = A - mu                                               # A
AB = list() ## Saltelli's Ci
AB[[1]] = forecastN(ICB,X,theta,alpha,Qmc,zp,Nmc) - mu   # ABi
AB[[2]] = forecastN(IC,XB,theta,alpha,Qmc,zp,Nmc) - mu   # ABd
AB[[3]] = forecastN(IC,X,thetaB,alphaB,QmcB,zp,Nmc) - mu # ABp
AB[[4]] = forecastN(IC,X,theta,alpha,Qmc,zpB,Nmc) - mu   # ABz
names(AB)=c("i","d","p","z")
#AB[[11]] = forecastN(IC,X,theta,alphaB,Qmc,zp,Nmc) - mu # ABa parameter uncertainty: random effects
#AB[[12]] = forecastN(IC,X,theta,alpha,QmcB,zp,Nmc) - mu # ABq parameter uncertainty about process error

## Saltelli eq 4.8: fij = E(Y|Xi,Xj) - fi - fj - E(Y) 
AB[[5]] = forecastN(ICB,XB,theta,alpha,Qmc,zp,Nmc) - mu   # ABid
AB[[6]] = forecastN(ICB,X,thetaB,alphaB,QmcB,zp,Nmc) - mu # ABip
AB[[7]] = forecastN(ICB,X,theta,alpha,Qmc,zpB,Nmc) - mu   # ABiz
AB[[8]] = forecastN(IC,XB,thetaB,alphaB,QmcB,zp,Nmc) - mu # ABdp
AB[[9]] = forecastN(IC,XB,theta,alpha,Qmc,zpB,Nmc) - mu   # ABdz
AB[[10]]= forecastN(IC,X,thetaB,alphaB,QmcB,zpB,Nmc) - mu # ABpz
names(AB)[5:10]=c("id","ip","iz","dp","dz","pz")
for(i in 5:10){
  AB[[i]] = AB[[i]] - AB[[substr(names(AB)[i],1,1)]] - AB[[substr(names(AB)[i],2,2)]]
}
```

## Decomposition at CATEGORY level
```{r}
d = length(AB)
f0f0 = apply(A*B,2,mean,na.rm=TRUE)
f02  = apply(A,2,mean,na.rm=TRUE)^2  ## Saltelli eqn 4.22
VE = EV = matrix(NA,d,NT)

## main effects: Si = V(fi)/V(Y)
for(i in 1:d){
  VEi = A*AB[[i]]   ## version from Saltelli eqn 4.21
  VE[i,] = apply(VEi,2,mean,na.rm=TRUE) - f0f0 ## Var(E(Y|Xi)); f0f0 from Rcpp
}
S = t(t(VE)/apply(VE,2, sum))

## interactions: Sij = Vij/V(Y)
## Saltelli 4.10c: Vij =V(fij(Xi,Xj))=V(E(Y|Xi,Xj))−V(E(Y|Xi))−V(E(Y|Xj))


## total effects: 
for(i in 1:d){
  EVi = B*AB[[i]]
  EV[i,] = apply(EVi,2,mean,na.rm=TRUE) - f0f0 ## E(Var(Y|Xi))
}
ST = t(1-t(EV)/apply(VE,2,sum))
```

```{r}
N.cols <- c("black","green","red","orange","cyan","blue") ## set colors
N.cols = RColorBrewer::brewer.pal(n=nrow(S),name = "Set3")
## main effects
Sc = apply(S,2,cumsum)
plot(time2,Sc[1,],ylim=c(0,1),type='n',main="Relative Variance: In-Sample",ylab="Proportion of Variance",xlab="time")
ciEnvelope(time2,rep(0,ncol(S)),Sc[1,],col=N.cols[1])
for(i in 2:nrow(S)){
  ciEnvelope(time2,Sc[i-1,],Sc[i,],col=N.cols[i])
}
legend("topleft",legend = rev(names(AB)),col=rev(N.cols),lty=1,lwd=5) #legend=c("Process","Q","RE","Parameter","Driver","InitCond")

```

## Todo: Q, random effects

